Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2015-12-11 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
4-core
Closes #73
NodIO-W name
Closes #80

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{baseline-times.png}
\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using {\sf NodEO}
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
characteristics: 
\begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with an 8-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) 
and Weibull (in red, left) for the {\sf NodIO$^2$} model.}  
\label{fig:gev:w2}
% File ../data/ips-puts.R
\end{figure}
%
As we did in the previous subsection, we have also fitted the number
of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown
in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
in the previous experiment, Table \ref{tab:puts:os}, we find higher values
for the location $\mu$ and scale $\sigma$ parameters, accounting for
the higher number of contributions per volunteer that has been
previously observed. However, the shape $\xi$ parameter is
substantially similar, with a {\em bell} shape leaning towards the
origin, indicating a pattern of a few users with many contributions
and many with few contributions. If we plot the model and the data
side by side, as shown in Figure \ref{fig:gev:w2} we see that the fit
is not so tight as in the previous case, with the model overestimating
the number of individuals with a high number contributions. This is
probably due to the fact that in the previous case, the user needed to
voluntarily reload the page to contribute the most generations, with a
{\em phase change} between the people that did so and the people that
just ran the experiment once until completion. In this case, however, the
user just needs to let it run, without that phase change and thus we obtain a
more {\em egalitarian} distribution of contributions, which rather
corresponds to a Weibull function, as shown in Figure
\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
equal to 0.69, a value that is remarkably similar to the 0.5 value
found for time devoted to games in Chambers et
al. \cite{chambers2005measurement}. It is important to note that, in
this case, the number of PUTs made per client does not correspond to a
uniform amount of computation, since 100 generations with a random
population belong, in each case, to different number of
operations. In other words, the same number of PUTs will correspond to
different number of operations in each case. However, since clients
are also different and take a different time in each case, it is
difficult to ascertain how this might have an influence in the
statistical distribution.

@@

\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO}-W$^2$ }
This experiment was deployed in a virtual private server with 512MB of memory
running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was expected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal approximations. In order to achieve this, the only change needed 
in the system was to limit the number of chromosomes kept in the pool. 
Previously, every PUT request which included a new chromosome increased 
the size of the pool, in a long-running experiment this has to be avoided
in order to not exhaust the available memory. In this experiment the size
of a chromosome was 1000 double precision numbers, considering this
and the memory available in the server, the the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant a random chromosome was pulled out of the pool before inserting a
new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
and {\tt classic-ea-float} modules were used. The parameters for the EA 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosome with the server every 100 generations. Again two workers in each window.
As before, a few participation calls were published in social networks.

>>>>

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{baseline-times.png}
\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using {\sf NodEO}
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
characteristics: 
\begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.49\linewidth]{gev-fit-ww.png}
\includegraphics[width=0.49\linewidth]{weibull-fit-ww.png}
\caption{Ranked number of {\tt PUT}s per worker and GEV fit (in red, right) 
and Weibull (in red, left) for the {\sf NodIO-W$^2$} model.}  
\label{fig:gev:w2}
% File ../data/ips-puts.R
\end{figure}
%
As we did in the previous subsection, we have also fitted the number
of HTTP {\tt PUT}s per worker to a  Generalized Extreme Value (GEV) function, with the result shown
in Table \ref{tab:puts:ww}. Comparing this table with what we obtained
in the previous experiment, Table \ref{tab:puts:os}, we find higher values
for the location $\mu$ and scale $\sigma$ parameters, accounting for
the higher number of contributions per volunteer that has been
previously observed. However, the shape $\xi$ parameter is
substantially similar, with a {\em bell} shape leaning towards the
origin, indicating a pattern of a few users with many contributions
and many with few contributions. If we plot the model and the data
side by side, as shown in Figure \ref{fig:gev:w2} we see that the fit
is not so tight as in the previous case, with the model overestimating
the number of individuals with a high number contributions. This is
probably due to the fact that in the previous case, the user needed to
voluntarily reload the page to contribute the most generations, with a
{\em phase change} between the people that did so and the people that
just ran the experiment once until completion. In this case, however, the
user just needs to let it run, without that phase change and thus we obtain a
more {\em egalitarian} distribution of contributions, which rather
corresponds to a Weibull function, as shown in Figure
\ref{fig:ipstime:w2} (right hand side). The fitted parameters for this
distribution, shown in Table \ref{tab:puts:ww}, show a shape parameter
equal to 0.69, a value that is remarkably similar to the 0.5 value
found for time devoted to games in Chambers et
al. \cite{chambers2005measurement}. It is important to note that, in
this case, the number of PUTs made per client does not correspond to a
uniform amount of computation, since 100 generations with a random
population belong, in each case, to different number of
operations. In other words, the same number of PUTs will correspond to
different number of operations in each case. However, since clients
are also different and take a different time in each case, it is
difficult to ascertain how this might have an influence in the
statistical distribution.

@@

\subsubsection{Setup of Rastrigin's experiment in {\sf NodIO-W$^2$ }
This experiment was deployed in a virtual private server with 512MB of memory
running Ubuntu OS version 14.04 and hosted in DigitalOcean.com. Unlike previous
experiments where the optimum was expected to be found many times, 
this problem was expected to run constantly for many days finding only
sub-optimal approximations. In order to achieve this, the only change needed 
in the system was to limit the number of chromosomes kept in the pool. 
Previously, every PUT request which included a new chromosome increased 
the size of the pool, in a long-running experiment this has to be avoided
in order to not exhaust the available memory. In this experiment the size
of a chromosome was 1000 double precision numbers, considering this
and the memory available in the server, the the pool was limited 
to 10,000 chromosomes.  In order to keep this size
constant a random chromosome was pulled out of the pool before inserting a
new one, when the pool was full. For this problem {\sf NodEO}'s {\tt chromosome-float} 
and {\tt classic-ea-float} modules were used. The parameters for the EA 
were tournament size = 3 and  population size = 500, returning exchanging a
chromosome with the server every 100 generations. Again two workers in each window.
As before, a few participation calls were published in social networks.
